Convex Surrogates via Polyhedral Losses

### 2019-02-01 cvx-fin.tex
working through the embedding via duality theorem

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}

@@

\begin{theorem}
  Let $\gamma:\simplex\toto\R$ be an elicitable finite property.
  Then $\gamma$ is $(n-1)$-embeddable, where $n=|\Y|$.
\end{theorem}
\begin{proof}
  Let $S:\R\to\reals^\Y$ be a scoring rule (negative loss) eliciting $\gamma$, and let $G:\simplex\to\reals$ be its expected score function (negative Bayes risk), given by $G(p) = \max_{r\in\R} p \cdot S(r)$.
  By Lemma~\ref{lem:finite-full-dim}, $G$ is polyhedral and therefore closed, and by~\citet[Theorem 19.2]{rockafellar1997convex}, its convex conjugate $C : \reals^\Y \to \reals$ is also polyhedral and closed.

  By \raf{results with Ian} we know that for some $\D \subseteq \partial G(\simplex) \subseteq \reals^\Y$ we have some bijection $\varphi : \R \to \D$ such that $\Gamma(p) = \varphi^{-1}(\D \cap \partial G(p))$.
  \raft{In other words, $\Gamma$ is a relabeling of subgradients of $G$}
  Thus, we can write
  \begin{equation}
    \label{eq:1}
    r \in \Gamma(p) \iff \varphi(r) \in \partial G(p) \iff p \in \partial C(\varphi(r))~,
  \end{equation}
  where the final equivalence follows from convex duality.
  \raft{Of course, we need to be careful here; this only holds if $G$ is proper and lower semi-continuous, but those should both hold for us.}

  Now define $L(\hat r,y) = C(\hat r) - \hat r \cdot \ones_y$ on the report space $\hat \R = \conv(\D)$;
  we will show that $L$ essentially elicits $\Gamma$.
  As $L$ is convex, we may write the optimality condition as
  \begin{equation}
    \label{eq:2}
    \hat r \in \Gamma_L(p) \iff \hat r \in \argmax_{\hat r'} \E_pL(\hat r',Y) \iff 0 \in \partial \E_pL(\hat r,Y)~,
  \end{equation}
  where of course the subgradient $\partial$ is with respect to $\hat r$.
  Plugging in our definition of $L$, we have
  \begin{align*}
    0 \in \partial \E_pL(\hat r,Y)
    & \iff 0 \in \partial \E_p \left(C(\hat r) - \hat r \cdot \ones_Y\right)
    \\
    & \iff 0 \in \partial \left(C(\hat r) - \hat r \cdot p\right)
    \\
    & \iff 0 \in \partial C(\hat r) - \{p\}
    \\
    & \iff p \in \partial C(\hat r)~.
  \end{align*}
  Putting this together with eq.~\eqref{eq:1}, we have for any $\hat r \in \D$ that
  \begin{equation}
    \label{eq:3}
    \hat r \in \Gamma_L(p) \iff p \in \partial C(\hat r) \iff \varphi^{-1}(\hat r) \in \Gamma(p)~,
  \end{equation}
  which implies that $L$ essentially elicits $\Gamma$ for this $\varphi$.
  (The trivially induced map $\varphi:\R\to\hat\R$ is injective as $\varphi:\R\to\D$ was a bijection.)
  \raf{If we assume $\Gamma$ is non-redundant, I believe we also have $\Gamma = \trim(\Gamma_L)$ using $\D$ is the report set.}
\end{proof}

>>>>

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\dom}{\mathrm{dom}}

@@

\raf{  By Lemma~\ref{lem:finite-full-dim}, $G$ is polyhedral and therefore closed, and by~\citet[Theorem 19.2]{rockafellar1997convex}, its convex conjugate $C : \reals^\Y \to \reals$ is also polyhedral and closed.
}

\begin{theorem}
  Let $\gamma:\simplex\toto\R$ be a non-redundant elicitable property, elicited by a loss $L:\R\to\reals^\Y_+$ whose Bayes risk is a closed function.
  Then $\gamma$ is $(n-1)$-embeddable, where $n=|\Y|$.
\end{theorem}
\begin{proof}
  Let $S:\R\to\reals^\Y$ be a scoring rule (negative loss) eliciting $\gamma$, and let $G:\simplex\to\reals$ be its expected score function (negative Bayes risk), given by $G(p) = \max_{r\in\R} p \cdot S(r)$.
  Let $C:\reals^\Y\to\reals\cup\{\infty\}$ be the convex conjugate of $G$, given by $C(u) = \sup_{p\in\simplex} p\cdot u - G(p)$.

  By~\citet[Theorem 2]{frongillo2014general} we know that for some $\D \subseteq \partial G(\simplex) \subseteq \reals^\Y$ we have a bijection $s : \R \to \D$ such that $\gamma(p) = s^{-1}(\D \cap \partial G(p))$.
  (In other words, $\gamma$ is a relabeling of subgradients of $G$.)
  Thus, we can write
  \begin{equation}
    \label{eq:1}
    r \in \gamma(p) \iff s(r) \in \partial G(p) \iff p \in \partial C(s(r))~,
  \end{equation}
  where the final equivalence follows from~\cite[Corollary 23.5.2]{rockafellar1997convex} as $G$ is a closed convex function.

  Now define $L:\dom(C)\to\reals^\Y$ by $L(u) = C(u)\ones - u$ where $\ones\in\reals^\Y$ is the all-ones vector.
  We will show that $L$ embeds $\gamma$.
  As $L$ is convex, we may write the optimality condition as
  \begin{equation}
    \label{eq:2}
    u \in \prop{L}(p) \iff u \in \argmin_{u'} p\cdot L(u') \iff 0 \in \partial p\cdot L(u)~,
  \end{equation}
  where the subgradient $\partial$ is with respect to $u$.
  Plugging in our definition of $L$, we have
  \begin{align*}
    0 \in \partial p\cdot L(u)
    & \iff 0 \in \partial p\cdot\left(C(u)\ones - u\right)
    \\
    % & \iff 0 \in \partial \left(C(u) - p\cdot u\right)
    % \\
    & \iff 0 \in \partial C(u) - \{p\}
    \\
    & \iff p \in \partial C(u)~.
  \end{align*}
  Putting this together with eq.~\eqref{eq:1}, and recalling that $s$ is a bijection, we have for any $u \in \D$ and $p\in\simplex$ that
  \begin{equation}
    \label{eq:3}
    u \in \prop{L}(p) \iff p \in \partial C(u) \iff s^{-1}(u) \in \gamma(p)~,
  \end{equation}
  which implies that $L$ embeds $\gamma$ with embedding function $\varphi:\R\to\reals^\Y$ given by $\varphi(r) = s(r)$.
  \raf{TO CHECK: the finiteness of $C$}
\end{proof}
